EoS-insensitive relations after GW170817

### 2019-01-15 paper/bin-love-after-gw170817.tex
new updates and cleaning to paper draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% Intro
The thermodynamic relationship between pressure and density (the equation of state, or EoS) of cold supranuclear matter, accessible in neutron stars, is critical to the study of such stars, and is one of the largest uncertainties in nuclear physics to date. 
% How to probe
The extraction of tidal deformabilities from the gravitational waveforms of binary neutron star merger events, such as GW170817, is a promising method of probing the nuclear structure.
% What was done before - Binary Love
Previous studies have shown that approximately EoS-insensitive ``binary Love universal relations" exist between symmetric and antisymmetric combinations of individual tidal deformabilities, which are difficult to be independently measured with second-generation gravitational wave interferometers.
% What was done before - I-Love-Q
Similarly, another set of EoS-insensitive relations exist between individual neutron star parameters: moment of inertia (I), tidal deformability (Love number), quadrupole moment (Q), and compactness (C) - known as the ``I-Love-Q" and ``C-Love" universal relations.
% What universal relations do
Such EoS-insensitive relations allow the elimination of some tidal parameters from the list of model parameters, thus breaking degeneracies and ultimately reducing the uncertainty of the dominant tidal effect in parameter estimation\kc{You mean $\tilde{\Lambda}$? In our paper with Carl and Aaron we showed that the binary Love relation doesn't affect the $\tilde{\Lambda}$ measurement much.}.
% What we will do
\kc{Add a sentence saying that even the more precise relations have some EoS variation which can cause a systematic error which can be marginalized over}
In this document, we explain how one can reduce equation of state variation for I-Love-Q, C-Love, and binary Love EoS-insensitive relations, which helps to reduce systematic errors on the tidal measurement of binary neutron star mergers.
However, we show that the systematic effects of these relations is very small until next-generation detectors. 
We achieve this  by restricting to only equations of state drawn from the 90\% posterior constraint on pressure as a function of density, as derived by the LIGO Collaboration.
% What we found
We find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with a mass ratio of 0.75, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim 50$\%.
% Conclude
We end by comparing systematic errors on tidal measurement due to the equation of state variation with statistical errors and comment on whether one can safely use these EoS-insensitive relations with future gravitational wave observations.
We conclude that the systematic errors injected from the use of improved EoS-insensitive relations become comparable to the statistical errors for future design upgrade Voyager.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
%Intro
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS up to the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy}, making NSs ideal laboratories for constraining ultra dense nuclear matter\kc{heavy ion collisions supposedly go to higher densities, but they are T-dependent}.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS.
For example, electromagnetic observations of the mass-radius relationship of NSs have been used to place limits on the NS EoS~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}.
However, these implications potentially suffer from large systematic errors due to astrophysical mismodeling uncertainties\kc{of...}.
Alternatively, the emission of gravitational waves (GWs) from binary NS merger events have proven to be a unique method of probing nuclear physics.
During the early inspiral portion of the merger, the orbital separation is large enough that the effect of companion tidal fields is negligible. 
As the separation decreases through GW emission, tidal forces are magnified and the NSs deform from sphericity, altering the orbital trajectories; a relic which is directly imprinted on the GW signal.
This deformation is characterized by the \textit{tidal deformability} $\Lambda$ as the linear response of the NSs quadrupole moment $Q_{ij}$ to the neighboring tidal field $\varepsilon_{ij}$~\cite{hinderer-love,Flanagan2008}.

%reparameterization of the template
In the case of binary NSs, each star becomes deformed via the process highlighted above, resulting in two highly correlated tidal parameters $\Lambda_1$ and $\Lambda_2$ entering in the gravitational waveform\cite{Flanagan2008,Vines:2011ud}.
Due to high correlations between between the two parameters, independent extraction becomes very difficult at current interferometer sensitivities~\cite{Wade:tidalCorrections}.
Typically, these high correlations can be reduced by strategically reparameterizing the waveform template with new tidal parameters constructed from linear combinations of $\Lambda_1$ and $\Lambda_2$.
For example, utilizing the parameters $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$~\cite{Favata:2013rwa,Wade:tidalCorrections} which enter the GW waveform at different post-Newtonian (PN) orders (powers of the ratio between the orbital speed and the speed of light $v/c^{2n}$) partially mitigates these correlations.
Similarly, the waveform may be parameterized by mass-independant tidal parameters $\lambda_0$ and $\lambda_1$, derived through a Taylor expansion about $m_0=1.4 \text{ M}_{\odot}$ of individual, unitful tidal deformability $\lambda \approx \lambda_0+\lambda+1(m_1-1.4 \text{ M}_{\odot})/\text{M}_{\odot}$~\cite{delPozzo:TaylorTidal}. 
Use of this set of tidal parameters allows one to effectively combine multiple events with varying mass and $\tilde\Lambda$ components, not possible in the former parameterization of $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$

@@

\subsection{Executive Summary}
\kc{The paper is fairly short, is the summary here necessary?}

%What we did
In this paper we attempt to improve I-Love-Q, C-Love, and binary Love EoS-insensitive relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love EoS-insensitive relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in EoS-insensitive relations.

@@

This is compared to the combined uncertainty $\sigma^A_N$ (turqoise shaded region) from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of expected binary NS merger detection rates for O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D over 1 observation year.

@@

The single-detection uncertainties are shown in unfilled blue circles, while the combined detection uncertaintied are outlined in dashed maroon.

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations\kc{cite reviews by Baym and Oertel}, one method to study a wide range of physically realizable EoSs is to parameterize them in a model-independent way\kc{how is this model-independant?}.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work~\cite{LIGO:posterior,Carney:2018sdv} sampled EoS parameters in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_0 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations\kc{cite}.
We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.

@@

\subsection{Neutron Star Tidal deformability}\label{tidal}
Binary neutron star mergers, such as GW170817, provide valuable insight into the internal structure of such stars. 
Namely, tidal effects which enter the gravitational waveform depend strongly on the EoS defining the NSs structure. 
The ($l=2$ electric type) \emph{tidal deformability}, denoted as $\lambda$, has the largest impact on the waveform phase. 
Consider a NS of mass $M$ in the presence of the tidal field $\varepsilon_{ij}$ of a companion star.
In response, the NS will deform away from sphericity under the acquisition of quadrupole moment $Q_{ij}$, characterized by $\lambda$ as the linear response to $\varepsilon_{ij}$~\cite{Flanagan2008,hinderer-love,Yagi2013}:
\begin{equation}
Q_{ij}=-\lambda \varepsilon_{ij}.
\end{equation}
One can extract $Q_{ij}$ and $\varepsilon_{ij}$ -- and thus $\lambda$, or its dimensionless form $\Lambda \equiv \lambda/M^5$ -- via diffent asymptotic limits of the gravitational potential:
\begin{align}
\begin{split}
\Phi=\frac{1+g_{tt}}{2}&=-\frac{M}{r} - \frac{3}{2}\frac{Q_{ij}}{r^3} \Bigg(\frac{x^i}{r} \frac{x^j}{r}-\frac{1}{3}\delta_{ij} \Bigg) + \mathcal{O} \Bigg( \frac{M^4}{r^4} \Bigg)\\
&+ \frac{1}{2} \varepsilon_{ij} x^i x^j + \mathcal{O} \Bigg( \frac{r^3}{M^3} \Bigg).
\end{split}
\end{align}
The metric component $g_{tt}$ is determined by first constructing a spherically symmetric, non-spinning background solution, followed by the introduction of a perturbative tidal deformation.
Next, the perturbed Einstein equations must be solved in the NSs interior given an EoS, then matched to the exterior Schwarzschild solution at the surface, modulo a constant.
This process is described in more detail in Ref.~\cite{hinderer-love}.
Further, the stellar mass and radius are determined from $p(R)=0$, and the above mentioned constant.
Similarly, the moment of inertia is obtained from the asymptotic behavior of the $g_{t \phi}$ metric component.

@@

\subsection{EoS insensitive relations}\label{sec:eosInsensitive}
Current gravitational wave interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, important previous work by Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations up to 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$, for a representative set of 11 EoSs. 
These relations, known as ``binary Love universal relations" allows one to break degeneracies between coupled tidal parameters.
This is important for two reasons: (i) allows us to algebraically eliminate tidal parameters from the template parameter list, improving parameter estimation, and (ii) automatic measurement of second tidal parameter given observation of the first.
It was shown through a simple Fisher analysis that the use of binary Love EoS-insensitive relations improved parameter extraction on $\tilde{\Lambda}$ by an order of magnitude.

@@

\subsubsection{Nuclear matter stars}\label{sec:ilq-nuc}
Following the work of Ref.~\cite{Yagi:ILQ}, the data for each EoS-insensitive relation is first fit to the following curve:
\begin{equation}\label{eq:ILQfit}
\ln{y_i}=a_i+b_i \ln{x_i} + c_i (\ln{x_i})^2 + d_i (\ln{x_i})^3 + e_i (\ln{x_i})^4,
\end{equation}
where the updated coefficients are given in Table~\ref{tab:ILQfit}.
Additionally, we offer an improvement to the functional form of the I-Love-Q fitting curve. 
Namely, we use the Newtonian relationships between various observables as a controlling factor in the fit~\cite{Yagi:ILQ}:
\begin{equation}\label{eq:Newtonian}
\bar{I}^{\text{N}} = C_{\bar{I}\Lambda}\Lambda^{2/5}, \hspace{3mm} \bar{Q}^{\text{N}} = C_{\bar{Q}\Lambda}\Lambda^{1/5}, \hspace{3mm} \bar{I}^{\text{N}} = C_{\bar{I}\bar{Q}}\bar{Q}^{2}.
\end{equation}
This is appended to an expansion in $\Lambda^{-1/5} \propto C$ (or $\bar{Q}^{-1/5}$), where $C=M/R$ is the compactness of the star.
This results in the following overall fitting relation:
\begin{equation}\label{eq:ILQfitNew}
y=C_{yx} x^{\alpha} \frac{1+\sum_{i=1}^3 a_i x^{-i/5}}{1+\sum_{i=1}^3 b_i x^{-i/5}},
\end{equation}
where $y$ and $x$ correspond to NS observables $\bar{I}$, $\bar{Q}$, and $\Lambda$, and $\alpha$ is given by $2/5$, $1/5$, and $2$ for the $\bar{I}-\Lambda$, $\bar{Q}-\Lambda$, and $\bar{I}-\bar{Q}$ relations, respectively.
The new fitting coefficients are presented in Table~\ref{tab:ILQfitNew}.
While the two above fits both result in fits with similar $R^2$ (Coefficient of determination, defined as $R^2=\sum_i(f_i-\bar{y})^2/\sum_i(y_i-\bar{y})^2$, where $\bar{y}$ is the mean data value, and $f_i$, $y_i$ are the modeled and actual data values) values of $\sim 0.9999995$ for the data presented, the lattter one has the advantage that it properly limits to the Newtonian case as $\Lambda \rightarrow \infty$~\cite{Yagi:binLove}.

@@

\subsection{C-Love Relations}\label{sec:clove}
In this section we consider approximate C-Love EoS-insensitive relations between NS compactness $\bar{C} \equiv M/R$ and unitless tidal deformability $\Lambda$, as discussed by Yagi and Yunes~\cite{Yagi:binLove}.
In addition, we show the R-Love relations, giving the relationship between NS radius and tidal deformability for the constrained and unconstrained sets of EoSs.
Similar to the previous section, we consider two classes of stars: nuclear matter EoSs and hybrid quark-hadron star EoSs.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% Intro
The thermodynamic relationship between pressure and density (the equation of state, or EoS) of cold supranuclear matter, accessible in neutron stars, is critical to the study of such stars, and is one of the largest uncertainties in nuclear physics to date. 
% How to probe
The extraction of tidal deformabilities from the gravitational waveforms of binary neutron star merger events, such as GW170817, is a promising method of probing the nuclear structure.
% What was done before - Binary Love
Previous studies have shown that approximately EoS-insensitive ``binary Love EoS insensitive relations" exist between symmetric and antisymmetric combinations of individual tidal deformabilities, which are difficult to be independently measured with second-generation gravitational wave interferometers.
% What was done before - I-Love-Q
Similarly, another set of EoS-insensitive relations exist between individual neutron star parameters: moment of inertia (I), tidal deformability (Love number), quadrupole moment (Q), and compactness (C) - known as the ``I-Love-Q" and ``C-Love" EoS insensitive relations.
% What universal relations do
Such EoS-insensitive relations allow the elimination of some tidal parameters from the list of model parameters, thus breaking degeneracies and ultimately reducing parameter extraction uncertainties.
% What we will do
However, even the most precise of EoS insensitive relations carry EoS variation which inject systematic errors into the parameter extraction process.
In this document, we explain how one can reduce equation of state variation for I-Love-Q, C-Love, and binary Love EoS-insensitive relations, which helps to reduce systematic errors on the tidal measurement of binary neutron star mergers.
However, we show that the systematic effects of these relations is very small until next-generation detectors. 
We achieve this  by restricting to only equations of state drawn from the 90\% posterior constraint on pressure as a function of density, as derived by the LIGO Collaboration.
% What we found
We find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with a mass ratio of 0.75, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim 50$\%.
% Conclude
We end by comparing systematic errors on tidal measurement due to the equation of state variation with statistical errors and comment on whether one can safely use these EoS-insensitive relations with future gravitational wave observations.
We conclude that the systematic errors injected from the use of improved EoS-insensitive relations become comparable to the statistical errors for future design upgrade Voyager.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
%Intro
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities \zc{Katerina do you know of any papers on this?}), making NSs ideal laboratories for constraining ultra dense nuclear matter.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS.
For example, electromagnetic observations of the mass-radius relationship of NSs have been used to place limits on the NS EoS~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}.
However, these implications potentially suffer from large systematic errors due to various x-ray burst astrophysical mismodeling uncertainties.
Alternatively, the emission of gravitational waves (GWs) from binary NS merger events have proven to be a unique method of probing nuclear physics.
During the early inspiral portion of the merger, the orbital separation is large enough that the effect of companion tidal fields is negligible. 
As the separation decreases through GW emission, tidal forces are magnified and the NSs deform from sphericity, altering the orbital trajectories; a relic which is directly imprinted on the GW signal.
This deformation is characterized by the \textit{tidal deformability} $\Lambda$ as the linear response of the NSs quadrupole moment $Q_{ij}$ to the neighboring tidal field $\varepsilon_{ij}$~\cite{hinderer-love,Flanagan2008}.

%reparameterization of the template
In the case of binary NSs, each star becomes deformed via the process highlighted above, resulting in two highly correlated tidal parameters $\Lambda_1$ and $\Lambda_2$ entering in the gravitational waveform\cite{Flanagan2008,Vines:2011ud}.
Due to high correlations between between the two parameters, independent extraction becomes very difficult at current interferometer sensitivities~\cite{Wade:tidalCorrections}.
Typically, these high correlations can be reduced by strategically reparameterizing the waveform template with new tidal parameters constructed from linear combinations of $\Lambda_1$ and $\Lambda_2$.
For example, utilizing the parameters $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$~\cite{Favata:2013rwa,Wade:tidalCorrections} which enter the GW waveform at different post-Newtonian (PN) orders (powers of the ratio between the orbital speed and the speed of light $v/c^{2n}$) partially mitigates these correlations.
Similarly, the waveform may be parameterized by mass-independent tidal parameters $\lambda_0$ and $\lambda_1$, derived through a Taylor expansion about $m_0=1.4 \text{ M}_{\odot}$ of individual, unitful tidal deformability $\lambda \approx \lambda_0+\lambda+1(m_1-1.4 \text{ M}_{\odot})/\text{M}_{\odot}$~\cite{delPozzo:TaylorTidal}. 
Use of this set of tidal parameters allows one to effectively combine multiple events with varying mass and $\tilde\Lambda$ components, not possible in the former parameterization of $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$

@@

\subsection{Executive Summary}
\zc{Me and Kent discussed whether or not we should include this section or not, and we decided like having the quick summary for busy readers - especially with the increasing number of figs and tables (8 tables, 14 figures so far). We can discuss this further if anyone doesn't want it included :)}
%What we did
In this paper we attempt to improve I-Love-Q, C-Love, and binary Love EoS-insensitive relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love EoS-insensitive relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in EoS-insensitive relations.

@@

This is compared to the combined uncertainty $\sigma^A_N$ (turquoise shaded region) from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of expected binary NS merger detection rates for O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D over 1 observation year.

@@

The single-detection uncertainties are shown in unfilled blue circles, while the combined detection uncertainties are outlined in dashed maroon.

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work~\cite{LIGO:posterior,Carney:2018sdv} sampled EoS parameters in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_0 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations~\cite{Zhao:massiveNS}.
We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.

@@

\subsection{Neutron Star Tidal deformability}\label{tidal}
Binary neutron star mergers, such as GW170817, provide valuable insight into the internal structure of such stars. 
Namely, tidal effects which enter the gravitational waveform depend strongly on the EoS defining the NSs structure. 
The ($l=2$ electric type) \emph{tidal deformability}, denoted as $\lambda$, has the largest impact on the waveform phase. 
Consider a NS of mass $M$ in the presence of the tidal field $\varepsilon_{ij}$ of a companion star.
In response, the NS will deform away from sphericity under the acquisition of quadrupole moment $Q_{ij}$, characterized by $\lambda$ as the linear response to $\varepsilon_{ij}$~\cite{Flanagan2008,hinderer-love,Yagi2013}:
\begin{equation}
Q_{ij}=-\lambda \varepsilon_{ij}.
\end{equation}
One can extract $Q_{ij}$ and $\varepsilon_{ij}$ -- and thus $\lambda$, or its dimensionless form $\Lambda \equiv \lambda/M^5$ -- via different asymptotic limits of the gravitational potential:
\begin{align}
\begin{split}
\Phi=\frac{1+g_{tt}}{2}&=-\frac{M}{r} - \frac{3}{2}\frac{Q_{ij}}{r^3} \Bigg(\frac{x^i}{r} \frac{x^j}{r}-\frac{1}{3}\delta_{ij} \Bigg) + \mathcal{O} \Bigg( \frac{M^4}{r^4} \Bigg)\\
&+ \frac{1}{2} \varepsilon_{ij} x^i x^j + \mathcal{O} \Bigg( \frac{r^3}{M^3} \Bigg).
\end{split}
\end{align}
The metric component $g_{tt}$ is determined by first constructing a spherically symmetric, non-spinning background solution, followed by the introduction of a perturbative tidal deformation.
Next, the perturbed Einstein equations must be solved in the NSs interior given an EoS, then matched to the exterior Schwarzschild solution at the surface, modulo a constant.
This process is described in more detail in Ref.~\cite{hinderer-love}.
Further, the stellar mass and radius are determined from $p(R)=0$, and the above mentioned constant.
Similarly, the moment of inertia is obtained from the asymptotic behavior of the $g_{t \phi}$ metric component.

@@

\subsection{EoS insensitive relations}\label{sec:eosInsensitive}
Current gravitational wave interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, important previous work by Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations up to 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$, for a representative set of 11 EoSs. 
These relations, known as ``binary Love EoS insensitive relations" allows one to break degeneracies between coupled tidal parameters.
This is important for two reasons: (i) allows us to algebraically eliminate tidal parameters from the template parameter list, improving parameter estimation, and (ii) automatic measurement of second tidal parameter given observation of the first.
It was shown through a simple Fisher analysis that the use of binary Love EoS-insensitive relations improved parameter extraction on $\tilde{\Lambda}$ by an order of magnitude.

@@

\subsubsection{Nuclear matter stars}\label{sec:ilq-nuc}
Following the work of Ref.~\cite{Yagi:ILQ}, the data for each EoS-insensitive relation is first fit to the following curve:
\begin{equation}\label{eq:ILQfit}
\ln{y_i}=a_i+b_i \ln{x_i} + c_i (\ln{x_i})^2 + d_i (\ln{x_i})^3 + e_i (\ln{x_i})^4,
\end{equation}
where the updated coefficients are given in Table~\ref{tab:ILQfit}.
Additionally, we offer an improvement to the functional form of the I-Love-Q fitting curve. 
Namely, we use the Newtonian relationships between various observables as a controlling factor in the fit~\cite{Yagi:ILQ}:
\begin{equation}\label{eq:Newtonian}
\bar{I}^{\text{N}} = K_{\bar{I}\Lambda}\Lambda^{2/5}, \hspace{3mm} \bar{Q}^{\text{N}} = K_{\bar{Q}\Lambda}\Lambda^{1/5}, \hspace{3mm} \bar{I}^{\text{N}} = K_{\bar{I}\bar{Q}}\bar{Q}^{2}.
\end{equation}
This is appended to an expansion in $\Lambda^{-1/5} \propto C$ (or $\bar{Q}^{-1/5}$), where $C=M/R$ is the compactness of the star.
This results in the following overall fitting relation:
\begin{equation}\label{eq:ILQfitNew}
y=K_{yx} x^{\alpha} \frac{1+\sum_{i=1}^3 a_i x^{-i/5}}{1+\sum_{i=1}^3 b_i x^{-i/5}},
\end{equation}
where $y$ and $x$ correspond to NS observables $\bar{I}$, $\bar{Q}$, and $\Lambda$, and $\alpha$ is given by $2/5$, $1/5$, and $2$ for the $\bar{I}-\Lambda$, $\bar{Q}-\Lambda$, and $\bar{I}-\bar{Q}$ relations, respectively.
The new fitting coefficients are presented in Table~\ref{tab:ILQfitNew}.
While the two above fits both result in fits with similar $R^2$ (Coefficient of determination, defined as $R^2=\sum_i(f_i-\bar{y})^2/\sum_i(y_i-\bar{y})^2$, where $\bar{y}$ is the mean data value, and $f_i$, $y_i$ are the modeled and actual data values) values of $\sim 0.9999995$ for the data presented, the lattter one has the advantage that it properly limits to the Newtonian case as $\Lambda \rightarrow \infty$~\cite{Yagi:binLove}.

@@

\subsection{C-Love Relations}\label{sec:clove}
In this section we consider approximate C-Love EoS-insensitive relations between NS compactness $C \equiv M/R$ and unitless tidal deformability $\Lambda$, as discussed by Yagi and Yunes~\cite{Yagi:binLove}.
In addition, we show the R-Love relations, giving the relationship between NS radius and tidal deformability for the constrained and unconstrained sets of EoSs.
Similar to the previous section, we consider two classes of stars: nuclear matter EoSs and hybrid quark-hadron star EoSs.

\subsubsection{Nuclear matter stars}\label{sec:clove-nuc}
Following Ref.~\cite{Yagi:binLove}, we begin by fitting the data for each set of EoSs to the simple curve:
\begin{equation}
C = \sum^2_{k=0} a_k (\ln{\Lambda})^k.
\end{equation}
Doing so yields $a_0 = 0.3616998$, $a_1 = -0.0354818$, and $a_2 = 0.0006193849$, similar to what was found in Ref.~\cite{Yagi:binLove}.
Additionally we offer an improvement to the fits as was done in Sec.~\ref{sec:ilq}. 
We consider the fitting function found in Eq.~\ref{eq:ILQfitNew}, with the controlling factor given by:
\begin{equation}
C^N=K_{C\Lambda}\Lambda^{-1/5}.\label{eq:cloveFit}
\end{equation}
The new fitting coefficients are displayed in Tab.~\ref{tab:ILQfitNew}, alongside the previously found I-Love-Q relations.
